---
title: 'Assignment 4: Rule management'
author: "Edwin de Jonge and Mark van der Loo"
date: "useR!2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(validatetools)
```


## Exercise 1, rule management

- What is maximum columns in a dataset you encountered in your work?
- Can you give an indication of the maximum number of data validity rules that are checked in production process ? 
- How many persons are involved in checking and maintaining the rules?


## Exercise 2 check boundaries

Load the R package `validatetools`.

a) Look at the file `code/rules.yml`

```yml
```{r, child="rules.yml", results='asis'}
```
```

and load the rules into R variable `rules` with the help of validator

```{r}
rules <- validator(.file = "rules.yml")
```


b) What are the allowed values for `age` and `income`?

c) Check this with `validatetools::detect_boundary_num`.


## Excercise 3, simplify

Simplify: 

a) 

```{r}
validator( if (income > 0) age >= 16
         , age < 12
        )
```
with `simplify_conditional`

b) 
```{r}
rules <- validator( r1 = if (income > 0) age >= 16
                  , r2 = age < 12
                  )
```

c) 
```{r, eval = FALSE}
rules <- validator( r1 = job %in% c("yes", "no")
                  , r2 = if (job == "yes") income > 0
                  , r3 = if (age < 16) income == 0
                  )
```
with Apply `simplify_rules(rules, job = "yes")`

d) Can you reproduce c) with the other simplifying functions?

## Excercise 4, find the conflicting rules

a) Open the file "infeasible_rules.txt" (e.g. `file.edit("infeasible_rules.txt")`).
Can you see which rules are in conflict?

b) Find which two rule(s) are causing the infeasibility in file "infeasible_rules.txt".
Look into the help file of `validatetools`.

```{r}
rules <- validator(.file = "infeasible_rules.txt")
is_infeasible(rules)
# do your thing...
```
