---
title: 'Assignment 3: data checking'
author: "Mark van der Loo and Edwin de Jonge"
date: "useR!2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1, variable checks

Variable checks are checks that can be performed on a field-by-field basis. For
example checking if `Age` is non-negative, or of integer type. Variable checks
are among the simplest checks.

We will use the `SBS2000` data set, that is included with `validate`.

Load the `SBS2000` data set.

```{r}
library(validate)
data(SBS2000)
```

a) Check the variable type 

The following `validator` defines a rule that column `turnover` should be a
`numeric`. When confronted with a data set (e.g. `SBS2000`) we see that this is
the case.

```{r}
rules <- validator(
  is.numeric(turnover)
)
out <- confront(SBS2000, rules)
summary(out)
```

Adjust the code, so it checks that `size` is a text variable and `staff` is a `integer`.
Explain why the size rule fails.

b) Create a rules object using `is.na()` to check missing items in `turnover` and `profit`. `confront` the rules with `SBS2000` and summarize the results.

c) Create a rules object using `field_length` (or `nchar`) to check that

-  `size` has at least 2 characters.
-  `id` codes have exactly 5 characters.
-  `size` has at least 2 and at most 3 characters. 

`confront` the rules set with these rules.

d) Check with the function `number_format` which of the following records has two digits.

```{r}
dat <- data.frame(x = c("2.54","2.66","8.142","23.53", "2.3"))
```

e) As d) but now with at least two digits.

## Exercise 2

The functions `is_complete()` and `all_complete()` are convenience functions
that test for missing values or combinations thereof in records. Have a look
at their help files.

a) 

- Create a rule set where you test ether `id` is complete.
- Create a rules set where you test ether the combination `id` and `turnover` 
is complete.
- `confront` the data with these rules.

```{r, echo=FALSE, eval=FALSE}
rules <- validator(
          is_complete(id)
        , is_complete(id, turnover)
)
out <- confront(SBS2000, rules)
summary(out)
```

b) Balance restrictions occur for example in economic micro data, where financial
balances must be met. Implement the following rules and check them on the data:

- profit is total revue minus total costs.
- total revenue is turnover plus other revenue
- profit is at most 60% of total revenue.

```{r, echo = FALSE, eval = FALSE}
rules <- validator(
    total.rev - profit == total.costs
  , turnover + other.rev == total.rev
  , profit <= 0.6*total.rev
)

out <- confront(SBS2000, rules)
summary(out)
```

c) `validate` also accepts conditional rules: `if {rule_p} rule_q`:
Execute the following:

```{r, eval=FALSE}
rule <- validator(if (staff >= 1) staff.costs >= 1)
out  <- confront(SBS2000, rule)
summary(out)
```

Note in the summary that the rule is rewritten: it turns the check into a 
vectorized check (which an `if` statement isn't), so it can be checked efficiently 
in a data.frame.

## Exercise 3

It is a good idea to store the validation rules apart from the data handling.
This make it more easy to reuse a set of validation rules in different parts 
of the process and even share the rules to others to communicate which 
quality checks have been done on the data, and a user can test if the data complies.

a) Put the following rules in a `rules.R` file

```r
turnover >= 0

staff >= 0

# profits can be negative (not for long...)
profit == total.rev - total.costs
```

and create a validator with `validator(.file="rules.R")`.

b) Rules can have metadata, this can be seen when turning a rule set into a data.frame

```{r, eval=FALSE}
rules_df <- as.data.frame(rules)
View(rules_df)
```

c) Use `names(rules)` to set the `id` of the rules to `BR01`, `BR02` and `BR03`.
You can do this in the same way as for renaming elements of a vector. Use the
`label` function to set the label of the third rule to "profit def". Use
`print(rules)` to see the differences.

d) Export the rule set to "br.yml" in yaml format using `export_yaml`.
Open "br.yml" in RStudio (or another text editor) and copy the comments of the rules
in `a)` into the descriptions of the yaml file.
Read the `br.yml` file into a validator object.

## Exercise 4

`validate` checks are normally executed in a data.frame. When your data is big, 
it is an option to execute the validation checks on a database using
the pkg `validatedb`.

These are the same checks as `validate` but translated into `SQL` and executed
on a data.base.

a) Execute the following code:

```{r,eval = FALSE}
library(validatedb)
# we are using a sqlite database in this demo
library(RSQLite)

# connnect with the database file
con <- DBI::dbConnect(SQLite(), "SBS2000.db")

# retrieve a handle to the table "enterprise" in this database file.
enterprise <- tbl(con, "enterprise")

print(enterprise)
```

b) Use confront to execute the rules on the database table and `print` and do a `summary`.


c) Use `values` on the result of confront see the contents of the checks.




