---
title: "Automating data quality measurement"
author: "Mark van der Loo and Edwin de Jonge"
date: "useR!2021"
output: 
  beamer_presentation:
    include:
      in_header: ../00tex/presheader.tex
classoption: "aspectratio=169"
---

## 

- Measuring differences in data
- Measuring evolution of quality


##

\begin{center}
\includegraphics[height=0.45\textheight]{fig/datastep}
\end{center}

##

\begin{center}
\includegraphics[height=0.7\textheight]{fig/datastep2}
\end{center}


##

\begin{center}
\includegraphics[height=0.6\textheight]{fig/cellsplit}
\end{center}


##

\begin{center}
\includegraphics[height=0.6\textheight]{fig/rulesplit}
\end{center}


## `lumberjack`

### Track changes in data

- Without _changing_ code
- Customizable, user-defined loggers
- Track multiple data sets
- Track a data set in multiple ways


## Main functions

\begin{center}
\begin{tabular}{ll}
\hline
Logger          & what it does \\
\hline
\texttt{start\_log}   & Assign a logger to an \texttt{R} object.\\
\texttt{stop\_log}    & Stop logging and dump log, where dumping can be switched off.\\
\texttt{dump\_log}    & Dump logging info and stop logging, where stopping is optional\\
\texttt{run\_file}    & Execute a file, while logging, in a  new \texttt{environment}.\\
\texttt{source\_file} & Execute a file, while logging, in the global \texttt{environment}.\\
\hline
\end{tabular}
\end{center}

## Built-in loggers

\begin{center}
\begin{tabular}{ll}
\hline
Logger          & what it does \\
\hline
\texttt{expression\_logger} & record result of custom \texttt{R} expressions.\\
\texttt{filedump}           & dump a file after each operation.\\
\texttt{simple}             & record whether anything changed (\texttt{logical}).\\
\texttt{cellwise}           & record cell-by-cell changes.\\
\hline
\end{tabular}
\end{center}

## Using lumberjack in a script

```{r, eval=FALSE}
library(lumberjack)
# create some loggers
logger1 <- filedump$new(dir="/my/output")
logger2 <- cellwise(key="id")
# add both loggers
start_log(dat, logger1)
start_log(dat, logger2)

# Code doing stuff with 'dat'

# dump one of the loggers
dump_log(dat, logger="filedump")

# continue modifying 'dat'
#...
```

## The expression logger

### Track any summary statistic

```{r, eval=FALSE}
logger <- expression_logger$new(
    av_turnover = mean(turnover, na.rm=TRUE)
  , av_staffcost = mean(staff.costs, na.rm=TRUE)
)

start_log(dat, logger)
```



## Create your own logger

### Loggers are reference (`R6`, or base `R`) objects with two obligatory methods

- `$add(meta, input, output)` Compute difference between in- and output
    - `meta`: R expression and source 
    - `input`: data before expression evaluation
    - `output`: data after expression evaluation
- `$dump()` Make the logger dump the logging info


## `validate` exports two loggers

### `lbj_cells()`

Keep track of diffence in data, using the `cells()` function.

```{r}
logger <- validate::lbj_cells()
```


### `lbj_rules`

Keep track of changes in rule violations, using the `compare()` function.

```{r, eval=FALSE}
logger <- validate::lbj_rules(rules)
```


## Exercise

1. Use the `lbj_rules()` logger to track changes in rule violations for the  `clean_supermarkets.R` script.
    - use the rules in `code/SBS2000_rules.R`
2. Add a logger that tracks the `mean` of variables `turnover` and `staff.costs`

### Note

- `R6` loggers are initialized with `$new()` (like the expression logger)
- Base R reference class loggers are initialized with `loggername()` (like `lbj_rules()`)



